---
title: "handsOn_dataWrangling"
author: "Kristine"
date: "2021-09-21"
output: html_document
---

# Part 2 - Data Wrangling

# Loading package

Write a code:
- loads "tidyverse" package
- if the package is not installed, what command you use to install it (add this line with # at the beginning to mark it as a comment)

```{r q1_P2}
install.packages("tidyverse")
#if you already have the package, you can just type library and it will show up on the tab button library("tidyverse")


```

Comment on:
- Why below code is not working (hint: here() is a function from "here" package)

```{r q2_P2, error=TRUE}

# did not load the correct function. we have to load the library
library(tidyverse)
install.packages("here")
library(here)
here()
```

Write a code:
- necessary to be able to use here() function above :)

```{r q3_P2}
install.packages("here")
library(here)
here()

```

# Loading data

- Create a folder named DATA under your project folder (RMED901)
- Download Konsultasjoner.csv on the github page under DATA folder

Write a code:
- loading data Konsultasjoner.csv on RStudio (remember to download data under "DATA" folder under your main project folder)
- open the data and look at it


```{r}
myData<-read_csv(here("DATA","Konsultasjoner.csv"))
```

#myData<-read_delim(here("DATA","Konsultasjoner.csv",delim="/t"))
```{r}
myData<-read_delim(here("DATA","Konsultasjoner.csv",delim="/t"))
```


#she added delim function to read the file differently. there were some strange observatoins. but i cant see it here because i dont have the here package available.
```{r}
myData%<%,pivot_longer(names_to="year",values_to="nConsultations",cols=4:12)
```
#if you want to change the columns, you cna use "separate" command. e.g you can choose to only have miminum and maximum age, instead of the whole age range. 
#r doesnt like numbers for objects or names on data. if you use numbers as an observation, use back ticks around it: ´1446`

