---
title: "Hands-on - Day 1"
author: "Kristine"
date: "2021-09-20"
output: html_document
---

# Part 1

# Calculation

Write a code:
- calculates "5 * 8"

```{r q1}
5*8
```

# Objects

Write a code:
- stores value "5" to object "a"
- stores value "8" to object "b"
- stores multiplication of "a" and "b" as "z"
- shows value of "z"

```{r q2}
a<-5
b<-8
z<-a*b
z
```

Comment on:
- Why below code is not working:

```{r q3, error =TRUE}
1a <- 5 # cannot start object name with a number
b <- 8
z <- 1a * b
z
```

# Functions

# R functions

Write a code:
- calculates square root of 10 by using an R function (hint: google square root in R)

```{r q4}
x<-10
sqrt(x)
```

Write a code:
- stores value of 3.14159265359 to object pi
- rounds value of pi to 2 decimals (hint: write "round" to search box on help tab and search for it or use ?round!)

```{r q5}
pi<-3.14159265359
round(pi,2)
```

Write a code:
- stores string "HeLLo" to object "word"
- shows "word" all lower case (hello) (hint: use tolower() function)
- shows "word" all upper case (HELLO) (hint: use toupper() function)

```{r q6}
word<-"HeLLo"
tolower(word)
toupper(word)
```

Comment on:
- Why below code is not working:

```{r q7, error =TRUE}
word2 <- HeLLo
tolower(word2)
toupper(word2)
# we didnt use double qoutes " so R doesnt identify it as an object. should be "HeLLo"
```

rnorm() function generates random variables from a normal distribution
Write a code:
- finds arguments of rnorm()
- comment on which arguments are optional 
- comment on default values of optional arguments

```{r q8}
args(rnorm)
# these arguments are optional: mean and sd
# default values of optional arguments are:0 for mean and 1 for sd
```

Write a code:
- generates 100 random normal values with a mean of 100 and sd of 15
- saves this to object named y
- checks mean and sd - are they 100 and 15 as we expect? (hint: use mean() and sd() functions)

```{r q9}
rnorm(100,mean=100,sd=15)
y<-rnorm(100,mean=100,sd=15)



```

You can also write your own functions!

Write a code:
- stores value "5" to object "a"
- stores value "8" to object "b"
- includes a function named "multiplication" to multiply two values
- calls this function with using "a" and "b" and storing the value as "z"
- shows value of z

```{r q10}
a<-5
b<-8
multiplication<-function(x,y){
res<-x*y
}
z<-multiplication(a,b)
z
```

Comment on:
- Why below code is not working:

```{r q11, error =TRUE}
a <- 5
b <- 8
multiplication <- function(){ 
  res <- x*y
}
z <- multiplication(a, b) 
z
# enter your comment here
```

# Vectors

Discuss and comment on:
- What is the type of y in q9? 
- What are the types of a and b in q2?
- What is the type of word in q6?
- Check their types by using class() and typeof() functions. Are there any differences?

Comments:

y is ...
a and b are ...
word is ...

```{r q12}



```

Write a code:
- Creates a vector with values 3, 6, 2 and 10, and stores this as "x" (hint: use function c(), get help by using ?c if you need)
- Calculates mean and sd values of x

```{r q13}



```

Comment on:
- Is the code below working as you expect?
- If it is not, why? (hint: check q13)

```{r q14, error=TRUE}
mean(3, 6, 2, 10) 
sd(3, 6, 2, 10)
# enter your comment here
```

Comment on:
- Is the code below working as you expect?
- If it is not, why?
- Can we calculate the mean for the numbers in this vector using mean() function? (hint: check arguments) Add this to the last line of code chunk.

```{r q15}
x <- c(3, 6, 2, NA)
mean(x)
# because of missing value NA, mean and sd calculations result in NA
mean(x,na.rm=T)
```

Write a code:
- shows 2nd and 4th elements of x

```{r q16}
x[2]
x[4]
40:50
y[40:50]
```

Comment:
- What happens when you run the following code:

```{r q17}
x <- c(3, 6, 2, NA)
x + 1
```

Comment on:
- the types of below vectors - is there anything unexpected?

```{r q18}
vec1 <- c(1, 7, 3)
typeof(vec1)

vec2 <- c("Height", "Weight", "BMI")
typeof(vec2)

vec3 <- c(1L, 7L, 3L) 
typeof(vec3)

vec4 <- c(T, F, TRUE)
typeof(vec4)

vec5 <- c(1, "BMI", 3)
typeof(vec5)

vec6 <- c("Height", "BMI", T)
typeof(vec6)

vec7 <- c(1, 7, NA)
typeof(vec7)
```

Comment on:
- Why below code is not working:

```{r q19, error =TRUE}
vec <- c(1, 4, "a")
vec + 1
```

Comment on:
- Why below code is not working:

```{r q20, error =TRUE}
vec1101 <- c(1, 4, 5)
result <- vec1011 + 1
vec1101+1
result
```

# End of Part 1




